WSClean: a new implementation of a fast, generic wide-field imager

### 2014-01-27 articles/2014/Article-WSClean/wsclean.tex
Progress

\section{The WSClean imager implementation} \label{sec:implementation}
We have implemented the $w$-stacking algorithm in the C++ language. Several steps are multi-threaded in this implementation: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and minor Clean iterations (peak-finding and image subtraction). For the latter, intrinsics were used as well. Because of these optimizations, minor Cleaning iterations are so fast that performing the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, is not so relevant. However, because the PSF varies for imaging with a non-zero w-value, major iterations are still required to fix the errors from using a constant PSF during minor iterations.

@@

\subsection{W-snapshots}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{MWA imaging example of the Vela and Puppis A supernova remnants. Left: normal projection centred on Puppis A, right: zenith projection to reduce $w$-values, but recentred on Puppis A. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
To get optimal imaging performance, the $w$-values can be kept small by phase-changing towards zenith. As shown, all-sky imaging can be quite efficient with this technique. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \cdot \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}. A recentred image can be stored in a Flexible Image Transport System (FITS) file, which allows setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. TODO show difference.

>>>>

\section{The WSClean imager implementation} \label{sec:implementation}
With the purpose of testing new algorithms for the imaging of MWA data, we have written a new imager around the $w$-stacking algorithm. Nevertheless, the imager is not specialized for the MWA, and has been succesfully used for other telescopes (TODO menton more).

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and minor Clean iterations (peak-finding and image subtraction). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations are so fast that performing the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, is not so relevant. However, the PSF varies for imaging with a non-zero w-value, and subtracting a constant PSF image space leads to inaccuracies. After a number of these minor iterations, it is therefore beneficial to invert the model back to the visibilities and subtract the model directly from the visibilities. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

\subsection{W-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, shifted towards Puppis A during imaging. Both images have been made with WSClean using the Cotton-Schwab clean algorithm. No beam correction was applied. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
To get optimal imaging performance, the $w$-values can be kept small by phase-changing towards zenith. As shown, all-sky imaging can be quite efficient with this technique. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \cdot \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. By doing the inverse corrections during visibility prediction, a ``shifted'' visibility set can be cleaned with major iterations similar to a non-shifted set.

A recentred image is in the projection of the zenith tangent plane, and will need an additional regridding step before multiple snapshots can be added together. An example of the difference in projection is given in Fig.~\ref{fig:vela-projection-example}, which displays a 13~min MWA observation imaged with WSClean. Cleaning an image that is in a different projection yields slightly different results, but qualitativily the two images are of equal accuracy. A recentred image can be stored in a Flexible Image Transport System (FITS) file, which allows setting the centre of tangent projection with the \texttt{CRPIXi} keyword \citep{wcs-in-fits}. Common viewers such as Kvis and DS9 support such FITS files and display their coordinates correctly. TODO show difference.
